Chemistry of fossil Pinus pollen and the effect of acetolysis

### 2021-03-09 manuscript.Rmd
added refs to intro

The study of plant micro and macrofossil (spores, seed and pollen) is an integral part of Palaeoecology and Palaeobotany.
Pollengrains are extracted from sediments, bogs and other deposits and counted.
These pollen records are an important tool to reconstruct past floristic, vegetational and environmental changes. 
EXAMPLES

@@

In recent years, the study of pollen grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrares (FTIR) spectroscopy has become a popular method for analysing pollen chemistry.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution.
Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods.
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen.
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes.

So far, most FTIR studies have used modern material or material extracted from herbarium specimens.
In some cases the pollen samples are processed with acetolysis following standard palynological methods [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989] to simulate fossilization.
The assumption is that acetolysis does not alter the chemistry of sporopollenin and is not affecting any chemical signal.
To date, there has not been a comparison between acetolysed and untreated fossil pollen to examine any effect of acetolysis on sporopollenin or chemical degradation of pollen during diagenesis.

>>>>

The study of plant micro and macrofossil (spores, seed and pollen) is an integral part of Palaeoecology and Palaeobotany.
Pollen records are an important tool to reconstruct past floristic, vegetational and environmental changes [@magriNewScenarioQuaternary2006; @tzedakisVegetationChangeGlacialInterglacial1994; @huntleyMigrationSpeciesResponse1989; @nogues-bravoCrackingCodeBiodiversity2018; @tinnerEcologyAbiesAlba2013].

@@

In recent years, the study of pollen grain chemistry has shown promising results for biological and palaeoecological research.
Fourier Transform Infrares (FTIR) spectroscopy has become a popular method for analysing pollen chemistry.
A number of FTIR studies have identified pollen chemical differences on genus level, which may extent to species level differences [@zimmermannChemicalCharacterizationIdentification2018; @jardineChemotaxonomyDomesticatedGrasses2019; @muthreichChemicalVariationsQuercus2020; @zimmermannInfraredSpectroscopyPollen2014; @depciuchIdentificationBirchPollen2018; @diehnDiscriminationGrassPollen2020]. 
This is a promising finding for the application of pollen identification with chemical methods in palaeoecological research, because it has the potential to increase taxomonic resolution @dellannaPollenDiscriminationClassification2009; @zimmermannAnalysisAllergenicPollen2016.
Further, chemical methods are suitable for automation and offer a more objective method than traditional morphological methods [ @mondolApplicationHighThroughputScreening2019; @zimmermannAnalysisAllergenicPollen2016; @guedesPollenRamanSpectra2014].
In addition to taxonomic differences in pollen chemistry, there is evidence of environmental signals that are recorded in the chemistry of pollen [@bagciogluMultiscaleVibrationalSpectroscopic2015; @kendelChemicalAnalysisPollen2020; @diehnCombiningChemicalInformation2020].
Several studies have shown links between solar and UV radiation and pollen chemistry as well as responses in pollen chemistry to temperature changes.

So far, most FTIR studies have used modern material or material extracted from herbarium specimens .
In some cases the pollen samples are processed with acetolysis following standard palynological methods [@erdtmanSixNotesPollen1959; @faegriTextbookPollenAnalysis1989] to simulate fossilization [@julierChemotaxonomyToolInterpreting2016; @woutersenNovelApproachStudy2018].
The assumption is that acetolysis does not alter the chemistry of sporopollenin and is not affecting any chemical signal.
To date, there has not been a comparison between acetolysed and untreated fossil pollen to examine any effect of acetolysis on sporopollenin or chemical degradation of pollen during diagenesis.
